### 2022-11-03 Scripts/additional-analysis.Rmd

```{r Original model CFA}
FF.model.Original <- "Observe =~ item1 + item6 + item11 + item15 + item20 + item26 + item31 + item36
             Describe =~ item2 + item7 + Ritem12 + Ritem16 + Ritem22 + item27 + item32 + item37
            Awareness =~ Ritem5 + Ritem8 + Ritem13 + Ritem18 + Ritem23 + Ritem28 + Ritem34 + Ritem38
            Nonjudge =~ Ritem3 + Ritem10 + Ritem14 + Ritem17 + Ritem25 + Ritem30+ Ritem35 + Ritem39
             Nonreact =~ item4 + item9 + item19 + item21 + item24 + item29 + item33 "

fit.original <- cfa(FF.model.Original, data = IRT, ordered = names(IRT),estimator = "WLSMV", mimic = "Mplus", test="standard")

lavaan::summary(fit.original, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (fit.original,c("gfi", "agfi", "nfi","rfi", "cfi.scaled","tli.scaled","rmsea", "srmr","aic"))

reliability(fit.original, return.total = TRUE)
```

```{r mod1, eval=FALSE, include=FALSE}
modfit.or <- modindices(fit.original, sort. = TRUE)
modfit.or[modfit.or$mi>3.84,]
```

```{r fulscale-refit-1}
#We accepted this model
##Previously Rtem 3 and Ritem 39 discarded due to loading lower than <.30
#Ritem38 is discared <.30
#Ritem17, Ritem12, 22, 16,35,30 is discarded for crossloading
FF.mod.cor.1 <- "Observe =~ item1 + item6 + item11 + item15 + item20 + item26 + item31 + item36
             Describe =~ item2 + item7 +   item27 + item32 + item37
            Awareness =~ Ritem5 + Ritem8 + Ritem13 + Ritem18 + Ritem23 + Ritem28 + Ritem34 
            Nonjudge =~  Ritem10 + Ritem14  + Ritem25 + Ritem30
             Nonreact =~ item4 + item9 + item19 + item21 + item24 + item29 + item33 
                                         "

fit.cor.mod.1 <- cfa(FF.mod.cor.1, data = IRT, ordered = names(IRT),estimator = "WLSMV", mimic = "Mplus")

lavaan::summary(fit.cor.mod.1, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (fit.cor.mod.1,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic"))

reliability(fit.cor.mod.1, return.total = TRUE)

```

```{r fullsacel-mod-1}
modfit <- modindices(fit.cor.mod, sort. = TRUE)
modfit[modfit$mi>3.84,]
```

```{r fulscale-refit-2}
#We are not reporting this model
##Previously Rtem 3 and Ritem 39 discarded due to loading lower than <.30
#Ritem38 is discared <.30
#Ritem17, Ritem12, 22, 16,35,30 is discarded for crossloading
FF.mod.cor.2 <- "Observe =~ item1 + item6 + item11 + item15 + item20 + item26 + item31 + item36
             Describe =~ item2 + item7 +   item27 + item32 + item37
            Awareness =~ Ritem5 + Ritem8 + Ritem13 + Ritem18 + Ritem23 + Ritem28 + Ritem34 
            Nonjudge =~  Ritem10 + Ritem14  + Ritem25 + Ritem30
             Nonreact =~ item4 + item9 + item19 + item21 + item24 + item29 + item33 
                                         item21 ~~  item24

                                         Ritem34 ~~  item33"

fit.cor.mod.2 <- cfa(FF.mod.cor.2, data = IRT, ordered = names(IRT),estimator = "WLSMV", mimic = "Mplus")

lavaan::summary(fit.cor.mod.2, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (fit.cor.mod.2,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic"))

reliability(fit.cor.mod.2, return.total = TRUE)

```

```{r}
compare <- anova(fit.cor.mod.1,fit.cor.mod.2)
summary(compare)
```

## Check Invariance####

```{r Shortcut Invariance}

library("semTools")
invariance.model <- measurementInvariance(model = FF.mod.cor.1,
                                          data = data,
                                          group = "Gender",
                                          ordered = names(data),
                                          estimator = "WLSMV",
                                          strict = T)
invariance.model[[1]] #configural Invariance
invariance.model[[2]] # loading, weak, metric
invariance.model[[3]] # intercept, strong, scaler
invariance.model[[4]] # residual, strict
invariance.model[[5]] # means, structural invariance
```

```{r Invariancedetail, include=FALSE}
# Configural invariance ####

configural <- cfa(model = FF.mod.cor.1,
                  data = data,
                  group = "Gender",
                 ordered = names(data),
                 estimator = "WLSMV")

summary(configural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (configural,c("gfi", "agfi", "nfi","rfi", 
                       "cfi","tli",
                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper","srmr"))

#Metric
weak <- cfa(model = FF.mod.cor.1,
                  data = data,
                  group = "Gender",
                  ordered =   names(data),
                  estimator = "WLSMV", 
            group.equal = "loadings")

summary(weak, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (weak,c("gfi", "agfi", "nfi","rfi", 
                       "cfi","tli",
                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper","srmr"))

First.comp <- compareFit (configural, weak)
summary(First.comp)

#Scaler
strong <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV", 
              group.equal = c("loadings", "intercepts"))
summary(strong, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

Second.comp <- compareFit  (weak, strong)
summary(Second.comp)

#residual
strict <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV",  
              group.equal = c("loadings", "intercepts", "residuals"))

summary(strict, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
fitmeasures (strict,c("gfi", "agfi", "nfi","rfi", "cfi.robust","tli.robust","rmsea", "srmr","aic"))

Third.comp <- compareFit(strict,strong)
summary(Third.comp)
#structural
structural <-cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV",  
                 group.equal = c("loadings", "intercepts", "residuals", "means","lv.variances","lv.covariances"))

summary(structural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
fitmeasures (structural,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic"))

fourth.comp <- compareFit  (structural,strict)
summary(fourth.comp)
comfit.par <- compareFit(configural, weak, strong, strict)
summary (comfit.par)

models <-  list("Configural" = configural, 
                "Metric" = weak, 
                "Scalar" = strong, 
                "Residual" = strict)

Invariance.table <- compareLavaan(models,
              nesting = "Configural > Metric > Scalar > Residual", 
              fitmeas = c("chisq.scaled", "df",  "cfi.scaled","tli.scaled","rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled" ),
              scaled = T,
              chidif = T, digits = 2)

colnames(Invariance.table) <- c("Chi-Square", "df", "CFI", "TLI", "RMSEA", "RMSEA 90% Lower CI", "RMSEA 90% Upper", "Chi-Square Difference", "df difference*", "p")

```

```{r InvarianceTab, results='asis'}
apa_table(Invariance.table, align = "c",  caption = "Invariance Analysis", note = " a = Metric vs Configural; b = Scalar vs Metric; c = Residual vs Scalar; d = Structural vs Residual;* =  df of model comparison", landscape = T, font_size = "footnotesize" )

```

```{r fulscale-refit-rejected, eval=FALSE, include=FALSE}
#Ritem 38 & 3,39 poor loading (<.3)
#item12 and  17, 22,16, 34 belonged to many factors

FF.model.Cor <- "Observe =~ item1 + item6 + item11 + item15 + item20 + item26 + item31 + item36
             Describe =~ item2 + item7 +   item27 + item32 + item37
            Awareness =~ Ritem5 + Ritem8 + Ritem13 + Ritem18 + Ritem23 + Ritem28 + Ritem34 
            Nonjudge =~  Ritem10 + Ritem14 + Ritem25 + Ritem30+ Ritem35
             Nonreact =~ item4 + item9 + item19 + item21 + item24 + item29 + item33
                                    Ritem23 ~~ Ritem34
                                    Ritem5 ~~  Ritem8
                                    Ritem28 ~~ Ritem34
                                   item36 ~~ Ritem35
                                    item9 ~~  item33
                                    item21 ~~  item24
  "

fit.cor <- cfa(FF.model.Cor, data = IRT, ordered = names(IRT),estimator = "WLSMV", mimic = "Mplus")
suma <- summary(fit.cor, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fit.indices.cor <- fitmeasures (fit.cor,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic"))

rel.cor <- reliability(fit.cor, return.total = TRUE)

```

```{r mod2, eval=FALSE, include=FALSE}
modfit.cor <- modindices(fit.cor, sort. = TRUE)
modfit.cor[modfit.cor$mi>3.84,]
```

```{r sec-order, include=FALSE}
#item 38 <.3
# item 10, 12, 16, 22, 14,25,30,35,19 load on several factors

FF.model.so <- "Observe =~ item1 + item6 + item11 + item15 + item20 + item26 + item31 + item 36
             Describe =~ item2 + item7 +  item27 + item32 + item37
            Awareness =~ Ritem5 + Ritem8 + Ritem13 + Ritem18 + Ritem23 + Ritem28 + Ritem34 
            Nonjudge =~ Ritem3 +   Ritem17 + Ritem39
             Nonreact =~ item4 + item9 +  item21 + item24 + item29 + item33 
Mindfulness =~ Observe + Describe + Awareness + Nonjudge +Nonreact 
Ritem5 ~~    Ritem8
item2 ~~   Ritem13
item21 ~~    item24
item11 ~~    item32
Ritem5 ~~   Ritem13
item9 ~~    item33
Ritem23 ~~   Ritem34
item6 ~~     item2
item36 ~~   Ritem17
"

fit.so <- cfa(FF.model.so, data = cfa.FSS, estimator = "MLR", mimic = "Mplus")
summary(fit.so, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (fit.so,c("gfi", "agfi", "nfi","rfi", "cfi.robust","tli.robust","rmsea", "srmr","aic"))

reliability(fit.so, return.total = TRUE)
```

```{r mod2, eval=FALSE, include=FALSE}
modfit.so <- modindices(fit.so, sort. = TRUE)
modfit.so[modfit.so$mi>3.84,]
```