### 2022-11-03 Scripts/simulation.Rmd

---
title: "Simulation"
output: html_document
---

```{r samplesizeIRT, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='hide', width=60}
library(SimDesign)
set.seed(222)
m <-  8 # item number
n <- c(100, 200, 300, 400, 450, 500, 550) # sample size
design <- as.data.frame(n)

irtGenerate <- function(condition, fixed_objects = FALSE) {
  n <- condition$n

  a <- matrix(rlnorm(m,.2,.3))
  diffs <- t(apply(matrix(runif(m*4, .3, 1), m), 1, cumsum))
  diffs <- -(diffs - rowMeans(diffs))
  diffs<- diffs + rnorm(m)
  d <- diffs
  dat <- simdata(a, d, n, itemtype = 'graded')
  return(dat)
}

irtAnalyze <- function(condition, dat, fixed_objects = NULL) {
  mod <- mirt(dat, 1, itemtype = 'graded', verbose = FALSE)
  simpars <- coef(mod, simplify = TRUE, digits = Inf)$items
  irtpars <- c(a = simpars[,1], d = simpars[,2])
  return(irtpars)
}

irtSummarize <- function(condition, results,
                         fixed_objects = NULL) {
  apop <- fixed_objects['alpha', ]
  dpop <- fixed_objects['d', ]
  simrmse <- RMSE(results, c(apop, dpop))
  out <- c(RMSE = simrmse)
  return(out)
}

simres <- runSimulation(design, replications = 100,
                        parallel = TRUE, generate = irtGenerate,
                        analyse = irtAnalyze, summarise = irtSummarize,
                        packages = c('mirt'))
simres

colind <- grep(".a.", colnames(simres))
sima <- as.data.frame(simres[, colind])
nvec <- as.numeric(simres$n)

meanRMSE <- rowMeans(sima)
names(meanRMSE) <- n
round(meanRMSE, 2)
matplot(nvec, log(sima), type = "l", col = 1, lty = 1,
        ylab = "log(RMSE)", xlab = "sample size",
        main = "Graded Response Model", xaxt = "n")
axis(1, at = nvec)
```